---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  

> Assuming the file is in current direcory  



```{r echo = TRUE}
   activity <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
  
````

> Use agrregate function to get the required data  
> Create histogram   

## What is mean total number of steps taken per day?  

```{r echo = TRUE}
        
      library(lattice)
      activity$date <- as.Date(activity$date, "%Y-%m-%d")
```   

> Create histogram    

```{r echo= TRUE}
        StepsperDay <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
        hist(StepsperDay$steps, main = "Total steps per day", xlab = "day", col = "Blue")
```
> Mean and the Median 

    
    ```{r echo = TRUE}
        mean(StepsperDay$steps)
        median(StepsperDay$steps)

    ```

## What is the average daily activity pattern?
> Make plot of interval (x axis) and the average number of steps taken, averaged across all
> days (y-axis)  


```{r echo = TRUE}
      activitypattern <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
      plot(row.names(activitypattern), activitypattern, type = "l", xlab = "5-minute interval", 
      ylab = "Average across all Days", main = "Average of daily Acitvity Pattern", 
     col = "orange")

```


>  Which 5-minute interval, on average across all the days in the dataset, contains the maximum >  number of steps?.      

``` {r echo= TRUE}
      max_steps <- which.max(activitypattern)
      names(max_steps)

````

## Imputing missing values.  

> the total number of rows with NAs  

    ``` {r echo= TRUE}

        activity_NA <- sum(is.na(activity))
        activity_NA
    ```
    
> Devise a strategy for filling in all of the missing values in the dataset.

> The strategy does not need to be sophisticated. For example, you could use the mean/median
> for  that day, or the mean for that 5-minute interval,etc    
> Missing data -Na replaced by mean in 5 min interval.    

``` {r echo= TRUE}
      AverageVal <- aggregate(steps ~ interval, data = activity, FUN = mean)
      fillNA <- numeric()
        for (i in 1:nrow(activity)) {
            obs <- activity[i, ]
            if (is.na(obs$steps)) {
                steps <- subset(AverageVal, interval == obs$interval)$steps
            } else {
              steps <- obs$steps
    }
      fillNA <- c(fillNA, steps)
}
      

```
> Missing Data -Create new dataset  with missing data filled with mean value.  


```{r echo= TRUE}
    new_data <- activity
    new_data$steps <- fillNA    

```
> Calculate and report the mean and median total number of steps taken per day.  

```{r echo= TRUE}

      RevisedStepsperDay<- aggregate(steps ~ date, data = new_data, sum, na.rm = TRUE)
```
> New The Histogram  

```{r echo= TRUE}

    hist(RevisedStepsperDay$steps, main = "Total steps per day", xlab = "day", col = "Blue")
```

> Revised Mean and Median.  


```{r echo= TRUE}
      mean(RevisedStepsperDay$steps)
      median(RevisedStepsperDay$steps)
```
## Are there differences in activity patterns between weekdays and weekend.  

> Create a new variable for indicating if a date falls on weekends or weekdays.  



```{r echo= TRUE}

    day <- weekdays(activity$date)
    dayIndicator <- vector()
    for (i in 1:nrow(activity)) {
      if (day[i] == "Saturday") {
          dayIndicator[i] <- "Weekend"
      } else if (day[i] == "Sunday") {
          dayIndicator[i] <- "Weekend"
      } else {
          dayIndicator[i] <- "Weekday"
      }
    }
        activity$dayIndicator <- dayIndicator
        activity$dayIndicator <- factor(activity$dayIndicator)

        dayPatterns <- aggregate(steps ~ interval + dayIndicator, data = activity, mean)
        names(dayPatterns) <- c("interval", "dayindicator", "steps")


```


> Panel plot containing a time series plot of the 5-minute interval (x-axis).
> and the average number of steps taken, averaged across all weekday days or weekend days.
> (y-axis).  



```{r echo= TRUE}
    
     xyplot(steps ~ interval | dayIndicator,dayPatterns, main="Average Steps per Day",type = "l",  xlab = "Interval", ylab = "Number of steps",layout = c(1, 2) )

``` 
